# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

1) Import raw datafile.

```{r}
setwd("~/Desktop/R/Coursera/ReproducibleResearch/RepData_PeerAssessment1")
ds = read.csv("activity.csv")
head(ds)
```

2)  (No general transformation necessary.)

## What is mean total number of steps taken per day?

1)  Make a histogram of the total number of steps taken each day

```{r}
ds.agg = aggregate(steps ~ date, data=ds, FUN=sum)
hist(ds.agg$steps, breaks=10)
```

2)  Calculate and report the mean and median total number of steps taken per day

```{r}
mean(ds.agg$steps)
median(ds.agg$steps)
```

## What is the average daily activity pattern?

1)  Make a time series plot of the average number of steps taken by interval.

```{r}
ds.int = aggregate(steps ~ interval, data=ds, FUN=mean)
plot(ds.int$steps, type="l")
````

2)  Interval 835 is the 5-minute interval contains the maximum number of average steps.

```{r}
int.max = max(ds.int$steps)
ds.int[which(ds.int$steps==int.max),]
```

## Imputing missing values

1) Calculate the number of missing values

```{r}
sum(is.na(ds$steps))
```

2-3)  Replace missing values for a date/interval with the mean for that interval.

```{r}
ds.imp = merge(ds,ds.int,by="interval",suffixes=c("",".int"),all.x=T)
ii = is.na(ds.imp$steps)
ds.imp$steps[ii] = ds.imp$steps.int[ii]
ds.imp = ds.imp[order(ds.imp$date,ds.imp$interval),]
head(ds.imp)
```

4)  Make a histogram of the total number of steps taken each day and calculate the mean/median per day.  Note that the mean is unchanged after the imputation and the median is similar.

```{r}
ds.imp.agg = aggregate(steps ~ date, data=ds.imp, FUN=sum)
hist(ds.imp.agg$steps)
mean(ds.imp.agg$steps)
median(ds.imp.agg$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

1)  Create a factor variable indicating whether the day is a weekday or weekend.

```{r}
ds.imp$weekday = weekdays(as.Date(ds.imp$date))
ii = ds.imp$weekday %in% c("Saturday","Sunday")
ds.imp$day.type[ii] = "weekend"
ds.imp$day.type[!ii] = "weekday"
ds.imp$day.type = factor(ds.imp$day.type)
str(ds.imp)
```


2) Plot time series of the average number of steps taken each interval by weekday/weekend.

```{r}
ds.weekday = aggregate(steps ~ interval + day.type, data=ds.imp, FUN=mean)
library(lattice)
xyplot(steps ~ interval | day.type, data=ds.weekday, type="l", layout=c(1,2))
```

